---
title: "Include Code Files"
author: Shafayet Khan Shafee
date: last-modified
format:
  html: 
    code-fold: true
    code-tools: true
engine: knitr
filters:
  - include-code-files
execute: 
  echo: true
---


> DISCLAIMER: The main focus of this extension is to enable `code-folding` for code chunk with included content from external files. `Show all code`, `Hide all code` from `code-tools` still would not work.

> Note: View the source code of this document by clicking `</> Code` on top-right corner.

You can include content from external file using either Pandoc Divs `:::` (easier) or using code chunk.


## Using Pandoc Divs

### `include-from`

To include codes using pandoc divs, use the `include-from` attribute to denote the path of the external file, relative to this qmd file. And within that Div, use a codeblock along with the name of the language of the content to be included to get correct syntax highlighting. (You can run `quarto pandoc --list-highlight-languages` to know for which languages, syntax highlighting is supported).


::: {include-from=test/snippets/HelloWorld.java}

```{.java}
```

:::

You can also pass options `start-line`, `end-line` to include a specific range of lines, `code-line-numbers` to enable source code line numbering, `code-filename` or `filename` to show a name of the file the included code is associated with.


### `start-line` & `end-line`

::: {include-from=test/snippets/HelloWorld.elm start-line=3 end-line=6 code-line-numbers=true}

```{.elm}
```

:::

Note that, we have used `code-line-numbers: true` to add line numbers in code chunk.

Also We can change the text `Code` to something nicer using `code-filename`.

## `filename` and `code-filename`

Using `filename` does not align properly with `code-folding`. Instead use `code-filename` when using `code-fold: true`.


::: {include-from=test/snippets/hello-world.adb code-line-numbers="true" code-filename=hello-world.adb}

```{.ada}
```

:::

**`code-filename` only works with `code-folding`. For othercases, use `filename`**



## Using Code chunk

Now there is a way to use code chunk instead of pandoc divs to include contents from external file. But there are two things to **note**,

- Since Quarto so far uses two types rendering engine, `knitr` and `jupyter`, we need to use code chunk accordingly.If you are rendering the quarto document using jupyter along with python kernel use {python} chunk to include code, similary if you are using julia kernel, you need to use {julia} and on the other hand if you are rendering the quarto document using knitr (i.e. engine: knitr) then use {r} chunk to include code, Or to avoid thinking about rendering engine, jupyter or knitr, you also can follow pandoc Div approach described above.

- You also need to use a chunk option `source-lang` to give the name of the language of the content to be included to get correct syntax highlighting.


> Note: View the source code of this document by clicking `</> Code` on top-right corner to see how code chunk is used.


## `include-from`

```{r}
#| include-from: test/snippets/capitalize.m
#| source-lang: matlab
#| code-filename: capitalize.m

# comment
```

**One very important detail to note that, you must put a random comment in code chunk so that the code chunk is not treated as empty, to make this filter work.**

## `start-line` & `end-line`

You can specify from which line to which line you want to include,


```{r}
#| include-from: test/snippets/capitalize.ts
#| source-lang: typescript
#| start-line: 2
#| end-line: 6
#| code-line-numbers: true
#| code-filename: capitalize.ts

# comment
```

Note that, we have used `code-line-numbers: true` to add line numbers in code chunk.


## Some more examples


```{r}
#| include-from: test/snippets/capitalize.php
#| source-lang: php
#| code-line-numbers: true
#| code-filename: capitalize.php

# comment
```


```{r}
#| include-from: test/snippets/fizz-buzz.rs
#| source-lang: rust
#| code-line-numbers: true
#| code-filename: fizz-buzz.rs

# comment
```


```{r}
#| include-from: test/snippets/fizz-buzz.jl
#| source-lang: julia
#| code-line-numbers: true
#| code-filename: fizz-buzz.jl

# comment
```
